\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{tabularx}

\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  citecolor=blue,
  urlcolor=blue
}
\setlength{\emergencystretch}{3em}
\allowdisplaybreaks

\newtheorem{assumption}{Assumption}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{observation}{Observation}

\newcommand{\stl}{\textsc{STL}}
\newcommand{\hal}{\mathrm{Hal}}
\newcommand{\baku}{\mathrm{Baku}}
\newcommand{\states}{\mathcal{S}}
\newcommand{\actions}{\mathcal{A}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\prob}{\mathbb{P}}
\newcommand{\valuefn}{V}
\newcommand{\simplex}{\Delta}
\newcommand{\val}{\operatorname{val}}
\newcommand{\gap}{\operatorname{gap}}

\title{From Exact Anchors to Simultaneous-Move AlphaZero\\
for the \stl{} Handkerchief Game}
\author{Repository whitepaper}
\date{July 15, 2026}

\begin{document}
\maketitle

\begin{abstract}
This whitepaper records both the implemented solver and the revised plan for
reinforcement learning. The game is modeled as a fully observed, two-player,
zero-sum stochastic game with simultaneous exact-second actions. The Python
engine is the rule oracle. The rigorous solver performs finite-horizon
backward induction with engine-derived chance and linear-programming minimax
at every simultaneous node. The approximate stack provides candidate search,
matrix-game Monte Carlo tree search (MCTS), tablebase frontiers, and a
policy/value network.

The repository does not yet implement a complete AlphaZero loop. Phases P0--P3
of the Python bridge now provide a finite episode contract, exact/search
conformance, reconstructable replay and checkpoints, and a tested
simultaneous-MCTS improvement object. The V4 static Generation-Zero ruler has
passed, but the first learned-MCTS audit mixed a horizon-one oracle with a
multi-step learned frontier. A horizon-aware V3 checkpoint contract, exact
V3-root/V2-successor Bellman closure, fresh sealing workflow, and aligned MCTS
audit are now implemented; their full generation and training runs remain
unexecuted. The compatibility self-play module still uses a legacy bucketed agent against
scripted opponents. This distinction matters: AlphaZero-style reinforcement
learning requires search policies at states actually visited by MCTS self-play
and terminal episode outcomes as value targets. Rust is deferred until the
Python behavior and evidence package are stable. We prove the finite-horizon exact result, the
non-expansiveness of minimax backup, a frontier-error bound, and a local saddle
gap bound. We also state precisely why these results do not prove convergence
or global exploitability of the learned agent.
\end{abstract}

\section{Scope, Status, and Claims}

The live runtime tree is under \path{stl/}. The disconnected old package mirror
was removed after pushed archival commit \texttt{a4b03db}; Git, rather than a
second source tree, retains that history. The implementation plan is
\path{stl/docs/REGEN2RL.md}, and the granular removal ledger is
\path{docs/DISTILLATION_AUDIT.md}. This paper supplies the mathematical
rationale and claim boundary.

\begin{center}
\begin{tabularx}{\textwidth}{@{}lX@{}}
\toprule
Path & Responsibility \\
\midrule
\path{stl/engine/} & Rule authority, clock, roles, legality, death, revival, and leap behavior \\
\path{stl/solver/exact.py} & Exact public state, engine expansion, LP minimax, and finite-horizon solve \\
\path{stl/solver/search.py} & Candidate search, bounded resolve, leaf adapters, and matrix-game MCTS \\
\path{stl/solver/tablebase.py} & Tactical fixtures, certified intervals, and Tier A lookup \\
\path{stl/learning/} & Targets, policy/value model, training, replay support, and gates \\
\path{stl/play/} & Playable agents, scripted opponents, environment wrappers, and legacy bucket baseline \\
\path{stl/commands/} & Hydra command implementations \\
\path{stl/config/} & Resolved experiment and command parameters \\
\bottomrule
\end{tabularx}
\end{center}

The following claims are justified now:

\begin{itemize}[leftmargin=*]
  \item engine-relative correctness of exact action legality and transition
        expansion, subject to the regression suite;
  \item exact minimax value for the finite-horizon game actually enumerated by
        the full-width LP solver;
  \item exact values of pinned analytic fixtures by their construction; and
  \item finite-depth best-response intervals under their reported frontier
        brackets.
\end{itemize}

The following claims are not justified now:

\begin{itemize}[leftmargin=*]
  \item that the unrestricted-duration game has been solved;
  \item that finite-budget candidate MCTS converges under the present
        selection rule;
  \item that held-out mean squared error implies low exploitability;
  \item that existing self-play code is AlphaZero self-play; or
  \item that a Rust implementation is required for, or already defines, the
        learning baseline.
\end{itemize}

\section{Game Model}

\subsection{Fully Observed State with Simultaneous Actions}

Let the players be \(\mathcal{N}=\{\hal,\baku\}\). At a nonterminal public
state \(s\in\states\), both players observe the complete state represented by
\path{ExactPublicState}: player cylinders, total time dead, death counts, life
status, physicality, referee CPR count, clock, half-round, round number, first
dropper, and terminal fields. The players then choose their current actions
without observing the other current action. This is a perfect-state or fully
observed simultaneous-move game, not an alternating-action board game and not
a hidden-state poker game.

The engine assigns a dropper \(D(s)\) and checker \(C(s)\). It is convenient
to write the two player action sets as \(A_{\hal}(s)\) and
\(A_{\baku}(s)\), then map the resulting pair to a drop second \(d\) and a
check second \(c\) according to the roles. The code stores matrices as
drop-by-check and transposes or negates them so that Hal remains the maximizing
player.

\subsection{Actor-Aware Exact-Second Legality}

Normal legal seconds are \(1,\ldots,60\). In the leap window only Baku, when
Baku is the dropper, may select second 61. Checkers are always capped at 60,
and Hal can never select 61. Thus
\[
A_D(s)=
\begin{cases}
\{1,\ldots,61\}, & D(s)=\baku \text{ and }s\text{ is a leap turn},\\
\{1,\ldots,60\}, & \text{otherwise},
\end{cases}
\qquad
A_C(s)=\{1,\ldots,60\}.
\]
Dense learned policies have length 62. Index 0 is illegal padding and index
\(j\) denotes literal second \(j\).

The check succeeds iff \(c\ge d\). On success the checker accumulates the
inclusive interval \(c-d+1\) seconds, including one second on the diagonal.
Literal second 1 is the immediate action; index 0 remains illegal padding. On failure the checker incurs
the fixed 60-second penalty. All death, revival, clock, and leap effects then
come from \path{stl/engine/game.py}.

\subsection{Transition Kernel and Utility}

Let
\[
P(s'\mid s,a_{\hal},a_{\baku})
\]
be the transition kernel induced by the engine, including survival and fatal
branches. The rigorous solver does not learn or reimplement this kernel.

For current death duration (q) and prior cumulative time dead (t), the
engine assigns zero revival probability when (q\ge300) or (t+q>300).
The second inequality is deliberately strict: resulting cumulative TTD exactly
300 remains eligible for revival. Cylinder capacity, current dose, and
cumulative TTD are therefore separate quantities even though they share the
same five-minute numerical boundary.

Terminal utility is always Hal-perspective:
\[
u(s)=
\begin{cases}
+1, & s\text{ is terminal and Hal wins},\\
-1, & s\text{ is terminal and Baku wins},\\
0, & s\text{ is an explicitly declared draw}.
\end{cases}
\]
There is no intermediate shaped reward in \path{stl/solver}.

\subsection{Why Reinforcement Learning Needs an Episode Contract}

\begin{observation}[The engine has no deterministic finite episode bound]
Every successful check now adds at least one second, so the old zero-ST
diagonal self-loop does not exist. A failed checker may nevertheless revive
and reset its cylinder repeatedly. The stochastic engine therefore does not
provide a deterministic half-round cap that is suitable as an RL batch
contract.
\end{observation}

Consequently, a completed-game outcome is not guaranteed merely by calling
the engine in a loop. The first Python RL baseline defines a capped training
game \(G_H\): after \(H\) half-rounds, a nonterminal trajectory is assigned
training result zero and marked as truncated. The initial plan freezes
\(H_{\mathrm{train}}=64\) and \(H_{\mathrm{eval}}=128\), then gates promotion
on low truncation rate and sensitivity to the cap. This is an RL-wrapper
adjudication, not a new engine terminal rule. Any result learned under this
contract is a result about \(G_H\), not automatically about the
unrestricted-duration game.

\section{Exact Finite-Horizon Solver}

\begin{assumption}[Engine authority]
The Python engine in \path{stl/engine} is the rule oracle. Solver correctness
is relative to its legality and transition functions.
\end{assumption}

Let \(b:\states\to[-1,1]\) be a boundary evaluator. For a terminal state,
the boundary is ignored in favor of \(u\). Define
\[
V_0^b(s)=
\begin{cases}
u(s), & s\text{ terminal},\\
b(s), & s\text{ nonterminal},
\end{cases}
\]
and, for \(h>0\), define the Hal-payoff matrix
\[
M_h^b(s)_{ij}
=
\sum_{s'}P(s'\mid s,a_{\hal,i},a_{\baku,j})
V_{h-1}^b(s').
\]
The Bellman-Shapley backup is
\[
V_h^b(s)=\val\!\left(M_h^b(s)\right)
=
\max_{\sigma\in\simplex(A_{\hal}(s))}
\min_{\tau\in\simplex(A_{\baku}(s))}
\sigma^{\mathsf T}M_h^b(s)\tau.
\]

For a matrix \(M\in\mathbb{R}^{m\times n}\), the row-player LP is
\[
\begin{aligned}
\max_{\sigma,v}\quad &v\\
\text{subject to}\quad
&\sum_{i=1}^{m}\sigma_i M_{ij}\ge v,
&&j=1,\ldots,n,\\
&\sum_{i=1}^{m}\sigma_i=1,\\
&\sigma_i\ge0,
&&i=1,\ldots,m.
\end{aligned}
\]

The certified code path uses full legal exact-second actions and this LP. With
the current terminal-only exact cutoff, \(b(s)=0\) and unresolved probability
is tracked separately.

\begin{proposition}[Exact finite-horizon value]
For fixed engine state \(s\), finite horizon \(h\), full legal action sets,
and boundary evaluator \(b\), recursive engine expansion followed by exact LP
minimax returns \(V_h^b(s)\).
\end{proposition}

\begin{proof}
The claim is immediate for \(h=0\). Assume it holds at \(h-1\). For each
legal joint action at \(s\), the engine enumerates all successor chance
branches and the induction hypothesis gives the continuation value at every
successor. Their probability-weighted sum is therefore the corresponding
entry of \(M_h^b(s)\). The one-step decision is a finite zero-sum matrix game,
so von Neumann's minimax theorem equates its LP value with
\(\val(M_h^b(s))\). This is \(V_h^b(s)\), completing the induction.
\end{proof}

\section{Error Reasoning for Depth-Limited Search}

\subsection{Minimax Backup Is Non-Expansive}

\begin{lemma}[Matrix-value Lipschitz bound]
If two equally shaped payoff matrices satisfy
\(\lVert M-\widehat M\rVert_{\infty}\le\epsilon\), then
\[
\left|\val(M)-\val(\widehat M)\right|\le\epsilon.
\]
\end{lemma}

\begin{proof}
For every pair of mixed strategies \((\sigma,\tau)\),
\[
\left|\sigma^{\mathsf T}(M-\widehat M)\tau\right|
\le\epsilon
\]
because \(\sigma\) and \(\tau\) are probability vectors. Taking a minimum
and then a maximum cannot enlarge this uniform bound. Applying the same
argument with \(M\) and \(\widehat M\) exchanged gives the absolute-value
claim.
\end{proof}

\begin{proposition}[Frontier error does not amplify under exact backup]
Suppose two boundary evaluators satisfy
\(\sup_s|b(s)-\widehat b(s)|\le\epsilon\). If all actions and transitions are
expanded exactly, then for every finite horizon \(h\),
\[
\sup_s|V_h^b(s)-V_h^{\widehat b}(s)|\le\epsilon.
\]
\end{proposition}

\begin{proof}
At depth zero the result is the hypothesis. If it holds at depth \(h-1\),
then every action-cell continuation differs by at most \(\epsilon\), since an
expectation is non-expansive in the supremum norm. The matrix-value lemma then
bounds the backed-up value difference by \(\epsilon\). Induction proves the
claim.
\end{proof}

This proposition justifies calibrated frontier values for an exact
depth-limited resolve. It does not say that a finite-sample MCTS estimate has a
uniform error bound, nor that a compact feature map can represent such a
bound.

\subsection{Local Policy Quality from Matrix Accuracy}

For a Hal row strategy \(\widehat\sigma\) and Baku column strategy
\(\widehat\tau\), define the exact matrix saddle gap
\[
\gap_M(\widehat\sigma,\widehat\tau)
=
\max_{\sigma}\sigma^{\mathsf T}M\widehat\tau
-
\min_{\tau}\widehat\sigma^{\mathsf T}M\tau.
\]
It is zero at an exact saddle point and measures the total one-step gain
available from unilateral deviations.

\begin{proposition}[Approximate-matrix saddle gap]
Let \((\widehat\sigma,\widehat\tau)\) be an exact equilibrium of
\(\widehat M\). If
\(\lVert M-\widehat M\rVert_{\infty}\le\epsilon\), then
\[
\gap_M(\widehat\sigma,\widehat\tau)\le2\epsilon.
\]
\end{proposition}

\begin{proof}
Let \(\widehat v=\val(\widehat M)\). Uniform matrix error implies
\[
\max_{\sigma}\sigma^{\mathsf T}M\widehat\tau
\le \widehat v+\epsilon,
\qquad
\min_{\tau}\widehat\sigma^{\mathsf T}M\tau
\ge \widehat v-\epsilon.
\]
Subtracting the second inequality from the first gives the result.
\end{proof}

This provides a concrete policy gate: whenever full-width exact matrices are
available, report saddle gap in addition to value error.

\subsection{Candidate Sets Need Their Own Audit}

No analogous bound follows merely because a candidate set contains many
actions or tactically plausible seconds. A restricted matrix can be identically
zero while one omitted Hal action pays \(+1\) against every Baku action. The
restricted value is then zero and the full value is one. Candidate search is
therefore an approximation whose omitted-action best-response gain must be
measured against full width on a frozen audit distribution.

\section{Current Search Stack}

\subsection{Selective Solve and Bounded Resolve}

\path{stl/solver/search.py} generates candidate seconds around endpoints,
diagonals, safe budgets, and overflow boundaries. Its selective recursion
still uses engine transitions and role-correct matrix minimax, but its action
restriction is approximate. Critical-state resolve applies a hard horizon,
usually one half-round, and evaluates the frontier with a tablebase or network
adapter. LP remains the certification path; Python CFR+ is a runtime option for
bounded local matrices.

The V5 Bellman smoke falsified the earlier special case that retained a compact
set when the checker had zero safe squandered seconds remaining. At one such
root the restricted value error was \(0.188889\) and the lifted full-width
saddle gap was \(0.222222\). Zero-safe-budget states therefore use the full
legal exact-second matrix; the same stored root then has zero value error and
saddle gap \(2.22\times10^{-16}\). This is a state-dependent exact fallback,
not a claim that the remaining candidate sets are certified without audit.

Bellman generation commits each root as a hash-bound bundle containing its
closure graph, replay rows, and exact-policy certificates. A deterministic
merge deduplicates successor states and remaps branch indices. Candidate-set
failure blocks sealing and training but cannot discard these exact artifacts;
only a passing gate can bind the concealed Bellman ruler into the one-use seal.

\subsection{Matrix-Game MCTS}

Each MCTS node stores candidate role actions, a Hal-perspective mean
\(Q\)-matrix, joint cell counts, priors, children, and an exact engine
snapshot. The current exploration term is
\[
U_{ij}=c_{\mathrm{puct}}P_{ij}
\frac{\sqrt{N}}{1+N_{ij}}.
\]
The Hal maximizer solves an optimistic matrix \(Q+U\); the opponent solves an
optimistic matrix for its own payoff, equivalently \(-Q+U\). Joint actions are
sampled from the two resulting role marginals, chance is sampled through the
engine, and Hal-perspective leaf values are backed up without alternating sign.

Let \(\widehat\sigma_D^{(t)}\) and
\(\widehat\sigma_C^{(t)}\) be role equilibria of the root's empirical
mean-\(Q\) matrix before the \(t\)-th root selection. The canonical improvement
policy is their linearly weighted average
\[
\pi_D=\frac{1}{W_T}\sum_{t=1}^{T}t\widehat\sigma_D^{(t)},
\qquad
\pi_C=\frac{1}{W_T}\sum_{t=1}^{T}t\widehat\sigma_C^{(t)},
\qquad
W_T=\sum_{t=1}^{T}t.
\]
Later empirical matrices receive greater weight without collapsing mixed play.
The optimism-augmented equilibria select exploratory joint cells but are not
reported as policy probability. Network priors are masked and normalized per
role, their outer product defines the joint prior, and optional root Dirichlet
noise perturbs the two legal marginals independently. The final saddle point
of the mean \(Q\)-matrix remains a separate value/policy diagnostic; in a
nearly flat matrix an LP may return an arbitrary pure vertex, so it is not the
canonical training or deployed policy.

The convergence theorem for simultaneous-move MCTS in
\cite{lisy2013} assumes Hannan-consistent selection, guaranteed exploration,
and a finite game tree. The repository's PUCT-like matrix selection, neural
frontier, candidate actions, transpositions, and finite episode wrapper are not
automatically covered by that theorem. The frozen exact-matrix convergence pack
therefore supplies empirical, not theorem-level, evidence: across 120
fixture/budget/seed records its 1024-iteration median absolute value error is
zero, 95th-percentile error is \(0.001812\), maximum saddle gap is
\(0.005726\), maximum per-fixture root-value standard deviation is
\(0.000736\), and no fixture worsens from 256 to 1024 iterations. A future
regret-matching SM-MCTS replacement remains appropriate if broader frozen
packs violate the declared gate.

\section{Audit of the Current Learning Attempt}

\subsection{Implemented Pieces}

The repository already contains several components needed by the target
system:

\begin{itemize}[leftmargin=*]
  \item \path{stl/learning/model.py}: a shared trunk, Hal value head, and two
        length-62 role-policy heads driven by 52 named V2 features;
  \item \path{stl/learning/contracts.py}: the versioned finite-episode,
        perspective, configuration, and horizon-sensitivity contract;
  \item \path{stl/learning/replay.py}: reconstructable
        \path{TrainingRecordV3} rows, safe hash-verified shards, collision
        audits, and grouped splits;
  \item \path{stl/learning/reachable.py}: legal engine trajectories whose
        snapshots can be replayed from the canonical opening before labeling;
  \item \path{stl/learning/train.py}: value MSE, masked role-policy cross
        entropy, source weights, grouped selection, and strict checkpoint
        bundles;
  \item \path{stl/learning/targets.py}: exact/tablebase labels and MCTS
        bootstrap labels;
  \item \path{stl/learning/support/reanalysis.py}: deeper exact attempts and a
        high-iteration MCTS fallback;
  \item calibration, audit, ladder, policy-drift, finite-depth best-response,
        SPRT, and promotion utilities; and
  \item \path{stl/play/agent.py}: a playable network-plus-MCTS agent that
        samples the canonical improved role policies.
\end{itemize}

The P1 and P3 frozen conformance packs test exact transition/orientation
parity, omitted-action gains, chance sampling, and horizon-one MCTS convergence.
The first P4 V2 smoke and full generation completed computationally but failed
the subsequent data audit. In the 458-row training shard, 370 nonterminal rows
violated the death/TTD/CPR reachability invariants; the external ruler had 215
such violations among 254 rows. All 48 terminal rows across both shards omitted
the terminal death counters, and 22 tactical exact states appeared in both
splits. Tier A was absent, while exact horizon and unresolved cutoff mass were
dropped during replay conversion. Those artifacts are therefore rejected
evidence, not a frozen Generation Zero.

\subsection{Blocking Integration Gaps}

\begin{itemize}[leftmargin=*]
  \item \textbf{No premature command surface.}
        The broken legacy \path{command=self_play} wrapper has been removed.
        A public command is added only when P5 supplies the joint MCTS actor,
        reconstructable replay, and command-contract tests.
  \item \textbf{Legacy actor.}
        \path{stl/learning/support/self_play.py} uses bucketed
        \path{CanonicalHal} against scripted opponents. It is curriculum or
        legacy outcome data, not MCTS self-play by both sides.
  \item \textbf{Expert-iteration labels.}
        MCTS bootstrap sweeps a constructed state grid and labels root value;
        it does not supply trajectory outcomes from the improved policy.
  \item \textbf{Generation Zero is not frozen.}
        The corrected V3 generator is implemented and structurally tested, but
        no corrected tiny or full corpus, external-ruler review, training run,
        or search-in-the-loop gate has run.
  \item \textbf{Incomplete promotion evidence.}
        Policy drift is optional in promotion, and finite-depth best response
        is called certified exploitability; the first omits a policy gate and
        the second can be misread as a global claim.
\end{itemize}

The correct response is not to discard the exact stack. It is to connect the
existing pieces with explicit schemas and gates, while quarantining legacy
bucket search from the new self-play actor.

\section{Target Python AlphaZero System}

\subsection{Reconstructable State and Replay}

Each \path{TrainingRecordV3} row carries a versioned 52-field feature vector
and a serialized exact public state. It also records both legal masks, both
improved role policies, Hal-perspective value target, target kind, episode and
half-round indices, truncation status, parent checkpoint digest, search/config
digest, exact value horizon, cutoff probability, value interval, state origin,
source artifact, legal action trace where applicable, schema versions, and RNG
seeds. Pickle-free numeric shards and a
canonical JSON manifest are written atomically and verified by hashes. This
turns replay into an auditable experiment record rather than an opaque tensor
dump.

The V2 encoder includes every nonconstant value-relevant field in
\path{ExactPublicState} and appends documented relative features. A
compact network remains an approximation, but replay reconstruction permits
collision analysis and exact or MCTS reanalysis. Split assignment is grouped
by exact-state hash, episode, and identical feature vector so repeated,
adjacent, or representationally collapsed states cannot cross train and
validation boundaries.

\subsection{Reachability and Isolation Certificates}

Let \(s_0\) be the canonical opening and let
\(T(s,d,c,\omega)\) be the engine transition for drop/check seconds and a
revival outcome. Define the reachable sets recursively by
\[
\mathcal{R}_0=\{s_0\},\qquad
\mathcal{R}_{k+1}=\left\{
T(s,d,c,\omega):s\in\mathcal{R}_k,\ d\in A_D(s),\ c\in A_C(s),
\ P(\omega\mid s,d,c)>0
\right\}.
\]
An engine-derived V3 row stores

\[
\tau=((d_1,c_1,\omega_1),\ldots,(d_k,c_k,\omega_k))
\]

and publication requires an independent replay to satisfy
\[
T^{\tau}(s_0)=s_{\mathrm{stored}}.
\]
At every death branch, forced survival requires \(p_{\mathrm{revive}}>0\), and
forced death requires \(p_{\mathrm{revive}}<1\). By induction on \(k\), the
base state is in \(\mathcal{R}_0\), and each certified step has legal actions
and positive-probability chance support; therefore every stored trajectory
snapshot lies in \(\mathcal{R}_k\). This is a constructive reachability
certificate, stronger than merely checking plausible counters.

Terminal rows are captured only after the engine applies the fatal transition.
Consequently the loser is dead, has at least one death and positive total
time-of-death, and the global CPR count equals the two players' total deaths.
Certified tactical and Tier A rows are not misrepresented as trajectory
snapshots; their distinct origin and source-artifact digest remain explicit.

For split isolation, construct a graph whose vertices are candidate rows and
whose edges connect equal exact-state hashes, equal episode IDs, or equal
feature hashes. Train/ruler assignment operates on connected components before
any LP label is computed. If \(C(s)\) denotes the component of row \(s\), then
\[
C(s_i)=C(s_j)\Longrightarrow
\operatorname{split}(s_i)=\operatorname{split}(s_j).
\]
Hence no connected state, trajectory, or feature-equivalence class can cross
the split boundary. A post-assembly audit recomputes all three intersections
and rejects publication unless each is empty.

Bellman roots require a stronger pre-label construction because one selected
root publishes labels for an entire one-transition family. Define

\[
 \Gamma(s)=\{s\}\cup
 \bigcup_{a\in A(s),\,b\in B(s)}
 \operatorname{supp}P(\cdot\mid s,a,b),
 \qquad
 \Gamma(R)=\bigcup_{s\in R}\Gamma(s).
\]

\begin{proposition}[Closure-separated split isolation]
Let (R_{\mathrm{tr}},R_{\mathrm{dev}},R_{\mathrm{ext}}) be Bellman root sets,
and let each published shard contain exactly the root and successor identities
in its corresponding (\Gamma(R)). Exact-state leakage between shards is absent
if and only if
\[
 \Gamma(R_i)\cap\Gamma(R_j)=\varnothing
 \quad\text{for every }i\neq j.
\]
Root disjointness (R_i\cap R_j=\varnothing) alone is not sufficient.
\end{proposition}
\begin{proof}
The exact-state identities published by shard (i) are precisely
(\Gamma(R_i)). Therefore a state is published by both shards (i) and (j)
exactly when it belongs to their intersection. Empty pairwise intersections are
thus sufficient, and any nonempty intersection supplies a leaking state, so
they are necessary. Since distinct roots can share a successor, disjoint root
sets need not have disjoint closures.
\end{proof}

The implemented allocator computes (\Gamma(s)) through engine transitions
before solving any (V_2) or (V_3) label, reserves the disclosed fixtures
first, then allocates the concealed ruler, development, and training roots
against all earlier reservations. The first root-only full allocation
demonstrated the necessity: its pairwise closure intersections contained
(95), (89), and (247) states. The corrected allocation contains
(5512), (2147), and (1874) reserved states with all three intersections
empty.

Finally, exact and interval labels have different semantics. An exact row stores
\[
(v,h,p_{\mathrm{cut}},[v,v]),
\]
where \(h\) is measured in half-rounds and \(p_{\mathrm{cut}}\) is unresolved
frontier probability. A Tier A row stores
\[
\left(\frac{\ell+u}{2},0,0,[\ell,u]\right),
\]
and its role-policy losses are inactive until a separate saddle certificate
exists. This prevents interval width from being mislabeled as probability and
prevents a one-ply heuristic policy from being advertised as exact.

\subsection{Policy/Value Function}

The network is
\[
f_{\theta}(x(s))=
\left(p_D(\cdot\mid s),p_C(\cdot\mid s),v_{\theta}(s)\right),
\]
where each policy is masked and normalized over the actor-aware legal seconds
for its role, and \(v_{\theta}(s)\in[-1,1]\) estimates Hal outcome. Role heads
are appropriate because the state encoder identifies which named player holds
each role.

The two policies define independent mixed actions. A joint correlated policy
would define a different strategic object, because it could coordinate the
players' simultaneous actions. Therefore the joint prior and sampling law are
factorized:
\[
P_{ij}=p_D(i\mid s)p_C(j\mid s),
\qquad
d\sim\pi_D(\cdot\mid s),
\quad
c\sim\pi_C(\cdot\mid s).
\]

\subsection{MCTS Self-Play}

At every visited state, one search produces both improved role marginals.
Self-play exploration perturbs the legal marginals separately and records the
noise seed. Evaluation disables noise. Because a simultaneous equilibrium may
require mixing, evaluation samples the noise-free marginals rather than taking
an argmax.

The actor stores temporary rows
\[
(s_t,\pi_{D,t},\pi_{C,t},m_{D,t},m_{C,t})
\]
and samples the next engine transition. At the end of the episode, it attaches
\(z\in\{-1,0,+1\}\) to every row, with zero reserved for the explicit capped
draw contract when no engine draw exists. Root MCTS value is logged for
diagnosis but is not mislabeled as the terminal outcome.

\subsection{Learning Objective}

For replay sample \((s,\pi_D,\pi_C,z)\), the AlphaZero-style loss adapted to
two role heads is
\[
\mathcal{L}(\theta)=
\lambda_v\left(z-v_{\theta}(s)\right)^2
-\lambda_D\sum_{a\in A_D(s)}\pi_D(a\mid s)\log p_D(a\mid s)
-\lambda_C\sum_{a\in A_C(s)}\pi_C(a\mid s)\log p_C(a\mid s)
+c\lVert\theta\rVert_2^2.
\]
This is the direct analogue of the AlphaZero objective
\cite{alphazero2017}, with separate simultaneous role policies. Exact and
tablebase anchors may substitute certified value and equilibrium-policy
targets. Rows without a certified policy have that policy loss disabled; they
do not receive a fabricated uniform label.

The initial replay mixture keeps 25 percent immutable exact/tablebase anchors
and 75 percent self-play rows. This ratio is an engineering baseline, not a
convergence theorem. Its purpose is to prevent catastrophic drift from exact
anchors while the empirical promotion gates judge whether self-play improves
strength.

\subsection{Reanalysis}

Reanalysis reconstructs stored states and refreshes the search policies using
the current frozen champion. Completed episode outcomes remain unchanged.
Exact and tablebase rows are immutable. Search-bootstrap values, if retained
for experiments, carry a distinct target-kind tag so they cannot be confused
with Monte Carlo outcomes.

\section{Promotion and Evidence}

Promotion is more conservative than AlphaZero's policy of continually using
the latest network. AlphaGo Zero's evaluate-and-select lineage, together with
the repository's observed calibration and strength regressions, motivates a
frozen champion/candidate gate.

Every candidate must pass:

\begin{enumerate}[leftmargin=*]
  \item artifact and configuration digest consistency;
  \item exact/tablebase value calibration and role-policy legality;
  \item the exact-matrix MCTS convergence pack;
  \item a paired two-direction arena that respects Hal/Baku asymmetry and uses
        a predeclared SPRT;
  \item non-regression against the scripted robustness league, including
        \path{pattern_reader};
  \item finite-depth best-response interval comparison at identical scenarios,
        horizons, support mass, and frontiers; and
  \item a mandatory policy-drift/readability report with exact-anchor saddle
        gaps.
\end{enumerate}

Best-response intervals certify only their bounded calculation. If
\([L_C,U_C]\) and \([L_B,U_B]\) overlap, the comparison is inconclusive. A
promotion report may say that no sampled case is certified worse; it may not
infer global improvement or global low exploitability.

\section{Phased Implementation and Rust Boundary}

The executable plan and numeric exit criteria live in
\path{stl/docs/REGEN2RL.md}. Its dependency chain is:

\begin{enumerate}[leftmargin=*]
  \item freeze the objective, perspective, and finite episode contract;
  \item certify engine/exact/search conformance in Python;
  \item introduce reconstructable replay and checkpoint schemas;
  \item validate matrix-game MCTS as the improvement operator;
  \item train a supervised exact-anchor generation zero;
  \item build genuine MCTS self-play for both roles;
  \item close replay, training, and reanalysis;
  \item integrate statistical promotion; and
  \item freeze a reproducible Python baseline and evidence package.
\end{enumerate}

Items 1--4 (P0--P3) are implemented and verified. Item 5 has produced and
consumed a successful V4 static holdout, but that horizon-blind checkpoint is
not admitted to search. The horizon-aware Bellman-closure implementation is
complete; its balanced exact generation, fixed candidate ladder, new one-use
holdout, and aligned MCTS evidence are the current stopping boundary.

No phase in this chain requires new Rust work. Existing Rust CFR+ bindings are
opt-in comparison fixtures only. After the Python baseline is frozen, narrow
profiled kernels may be ported behind parity tests. Python remains the oracle
for legal actions, transitions, policy normalization, replay semantics, and
promotion. The Rust acceptance ordering is behavioral parity first and speed
second.

\section{What the Proofs Do and Do Not Establish}

The exact induction proves a finite-horizon result when all legal actions and
chance branches are represented and the LP is solved. The non-expansive backup
and saddle-gap propositions explain how a genuine uniform frontier or matrix
error bound would control local value and policy error.

They do not provide such a uniform error bound for a neural network. Empirical
MSE is distribution-dependent; feature collisions can force incompatible
states to share a prediction. Finite-budget MCTS adds sampling and selection
error. Candidate sets add unmeasured action error unless full-width audits are
performed. Function approximation, replay nonstationarity, and policy/value
co-adaptation remove the assumptions of tabular exact policy iteration.

Accordingly, the planned system is a rigorously anchored empirical solver, not
a theorem of global AlphaZero convergence. A future global claim would require
at least a well-defined unrestricted objective, exhaustive or otherwise valid
full-game bounds, and a global exploitability certificate. Until then, reports
must keep these categories separate:

\begin{itemize}[leftmargin=*]
  \item \textbf{exact}: full-width finite-horizon LP or analytic tablebase;
  \item \textbf{bounded}: interval or error statement for a declared horizon
        and frontier;
  \item \textbf{approximate}: finite-budget candidate MCTS or neural output;
  \item \textbf{empirical}: self-play, arena, calibration, and robustness
        results on declared distributions.
\end{itemize}

\section{Generation-Zero Generalization Firewall}

The first learned evaluator is admitted to search only after a three-way
development protocol.  A previously queried ruler is diagnostic evidence and
cannot remain a holdout.  The replacement benchmark is generated and sealed
before optimization; its digest is bound to exactly one checkpoint and gate
configuration when evaluation begins.

For a boundary state whose current checker has cylinder accumulation \(300\),
every legal joint action injects a fatal dose into that checker.  Hence the
continuation value is invariant to timing but not to role order:
\[
V(s)=\begin{cases}
+1,&\text{Baku is the checker},\\
-1,&\text{Hal is the checker}.
\end{cases}
\]
The benchmark crosses this invariant with single and double overflow, clock
class, and physically valid death/TTD histories.  In the Leap Second interior
construction, Baku's legal drop at second \(61\) defeats every Hal check in
seconds \(1\) through \(60\).  For dose \(q=x_H+60<300\),
\[
V(s)=p_{\mathrm{rev}}(q,\eta_H,n_R)(+1)
 +(1-p_{\mathrm{rev}}(q,\eta_H,n_R))(-1)
 =2p_{\mathrm{rev}}-1,
\]
where \(\eta_H\) contains Hal's own death and TTD history and \(n_R\) is the
referee CPR count.  Varying only the opponent's history is therefore not a
valid generalization test.

Aggregate MSE is insufficient because many easy invariant states can hide one
missing interaction.  Promotion consequently requires source MSE, semantic
family MSE, and maximum per-row error simultaneously.  Model correction first
freezes the shared trunk and policies; trunk updates are permitted only after
that conservative stage fails development, and any such update requires a new
certificate-backed policy audit.  No MCTS result is promotion evidence until
the one-use full holdout passes.

The implemented V4 run makes the distinction concrete.  The sealed split
contained $4483$ training, $871$ development, and $849$ holdout rows,
with empty pairwise intersections of exact-state, episode, and feature hashes.
The selected model passed its one-use holdout with overall MSE $0.002281$,
boundary/interior tablebase MSE $0.000177/0.000838$, and maximum exact-matrix
saddle gap $0.028476$.  These results certify interpolation on the declared
holdout distribution; they do not certify its use as a recursive leaf oracle.

To see why, let $D_0$ denote the sealed root distribution and let
$D_1^{\pi}$ be the successor distribution induced by an MCTS sampling policy
$\pi$.  The learned cell estimate is
\[
  \widehat Q(s,a,b)
    = \mathbb{E}_{S'\sim P(\cdot\mid s,a,b)}[\widehat V(S')].
\]
Consequently,
\[
  |\widehat Q(s,a,b)-Q(s,a,b)|
  \leq
  \mathbb{E}_{S'\sim P(\cdot\mid s,a,b)}
  |\widehat V(S')-V(S')|.
\]
A small $\mathbb{E}_{s\sim D_0}(\widehat V(s)-V(s))^2$ supplies no bound on
the right-hand side when $D_1^{\pi}$ is not covered by $D_0$.  This is the
successor-closure requirement: every state distribution on which search asks
the network for a value needs exact development evidence or a separately
validated error bound.

The first candidate-guided MCTS audit reported median, 95th-percentile, and
maximum absolute root-value errors $0.183849$, $0.910766$, and $0.911683$ at
$1024$ simulations, with maximum full-width saddle gap $0.804233$.  The
configuration is rejected.  It does not, however, isolate successor
generalization: the audit's oracle was terminal-frontier $V_1$, whereas the
candidate supplied longer-horizon continuation at depth-one leaves.  On the
two safe fixtures the exact values are $V_1=0/0.5$ but approximately
$V_3=0.737/0.75$.  Comparing those objects is not a valid evaluator-error test.

The repaired contract writes the approximation as $\widehat V(s,h)$, with
the finite horizon supplied separately from the physical state.  For the first
bounded learned search, root horizon is three and a single transition leaves
horizon two:
\[
 \widehat Q_3(s,a,b)
 =\sum_{s'}P(s'\mid s,a,b)\widehat V(s',2),\qquad
 \widehat B_3(s)=\val(\widehat Q_3(s)).
\]
The development gate measures both accuracy
$|\widehat B_3(s)-V_3(s)|$ and internal consistency
$|\widehat V(s,3)-\widehat B_3(s)|$.  The second quantity alone is not an
accuracy certificate: two jointly biased predictions may be perfectly
consistent.

\begin{proposition}[Bellman-closure error bound]
For a stored root whose full legal matrix and all engine chance branches are
present,
\[
 |\widehat B_3(s)-V_3(s)|
 \leq \|\widehat Q_3(s)-Q_3(s)\|_\infty
 \leq \max_{a,b}\sum_{s'}P(s'\mid s,a,b)
       |\widehat V(s',2)-V_2(s')|.
\]
\end{proposition}
\begin{proof}
The first inequality is non-expansiveness of the zero-sum matrix value under
the sup norm.  For each cell, linearity of chance expectation followed by the
triangle inequality bounds its error by expected successor error.  Taking the
largest legal cell gives the result.
\end{proof}

The V5 artifact therefore stores every legal cell, every positive-probability
engine branch, and every unique exact $V_2$ successor, and verifies
$V_3(s)=\val(Q_3(s))$ before training.  A new sealed holdout binds this closure
to fresh calibration evidence.  Self-play remains unauthorized until one
checkpoint passes direct, backed, residual, saddle-gap, and aligned MCTS gates.

\section{Validation Protocol}

The minimum documentation-aligned regression commands are:

\begin{verbatim}
uv run pytest tests/engine -q
uv run pytest tests/solver -q
uv run pytest tests/learning -q
uv run pytest tests/play -q
uv run pytest -q
uv run pytest --collect-only -q
\end{verbatim}

The stable-baseline gate additionally requires deterministic self-play replay,
exact-state reconstruction, grouped split isolation, exact-matrix MCTS budget
curves, a successful and a rejected promotion dry-run, horizon sensitivity,
and a graph refresh by running \texttt{graphify update} against the repository
root; generated
checkpoints and reports remain outside source review unless the artifact itself
is under audit.

\section{Conclusion}

The repository has a credible exact/search foundation: engine authority,
literal-second stochastic minimax, tablebase anchors, bounded resolve, a tested
matrix-game MCTS improvement object, reconstructable replay/checkpoints, and a
policy/value learner. The remaining gap is a frozen Generation Zero, genuine
MCTS self-play trajectories with terminal outcomes, a closed replay/reanalysis
loop, and mandatory statistical promotion. Only after those Python phases
produce a stable evidence package is the baseline suitable as an oracle for a
later Rust kernel.

\begin{thebibliography}{99}

\bibitem{vonneumann1928}
John von Neumann.
\newblock Zur Theorie der Gesellschaftsspiele.
\newblock \emph{Mathematische Annalen}, 100:295--320, 1928.
\newblock \url{https://doi.org/10.1007/BF01448847}

\bibitem{shapley1953}
Lloyd S. Shapley.
\newblock Stochastic games.
\newblock \emph{Proceedings of the National Academy of Sciences},
39(10):1095--1100, 1953.
\newblock \url{https://doi.org/10.1073/pnas.39.10.1095}

\bibitem{zinkevich2007}
Martin Zinkevich, Michael Johanson, Michael Bowling, and Carmelo Piccione.
\newblock Regret minimization in games with incomplete information.
\newblock \emph{Advances in Neural Information Processing Systems}, 2007.
\newblock \url{https://papers.nips.cc/paper_files/paper/2007/hash/08d98638c6fcd194a4b1e6992063e944-Abstract.html}

\bibitem{lisy2013}
Viliam Lisy, Vojtech Kovarik, Marc Lanctot, and Branislav Bosansky.
\newblock Convergence of Monte Carlo tree search in simultaneous move games.
\newblock arXiv:1310.8613, 2013.
\newblock \url{https://arxiv.org/abs/1310.8613}

\bibitem{deepstack2017}
Matej Moravcik, Martin Schmid, Neil Burch, Viliam Lisy, Dustin Morrill,
Nolan Bard, Trevor Davis, Kevin Waugh, Michael Johanson, and Michael Bowling.
\newblock DeepStack: Expert-level artificial intelligence in heads-up
no-limit poker.
\newblock \emph{Science}, 356(6337):508--513, 2017.
\newblock \url{https://doi.org/10.1126/science.aam6960}

\bibitem{alphagozero2017}
David Silver, Julian Schrittwieser, Karen Simonyan, Ioannis Antonoglou,
Aja Huang, Arthur Guez, Thomas Hubert, Lucas Baker, Matthew Lai, Adrian Bolton,
Yutian Chen, Timothy Lillicrap, Fan Hui, Laurent Sifre,
George van den Driessche, Thore Graepel, and Demis Hassabis.
\newblock Mastering the game of Go without human knowledge.
\newblock \emph{Nature}, 550:354--359, 2017.
\newblock \url{https://doi.org/10.1038/nature24270}

\bibitem{alphazero2017}
David Silver, Thomas Hubert, Julian Schrittwieser, Ioannis Antonoglou,
Matthew Lai, Arthur Guez, Marc Lanctot, Laurent Sifre, Dharshan Kumaran,
Thore Graepel, Timothy Lillicrap, Karen Simonyan, and Demis Hassabis.
\newblock Mastering chess and shogi by self-play with a general reinforcement
learning algorithm.
\newblock arXiv:1712.01815, 2017.
\newblock \url{https://arxiv.org/abs/1712.01815}

\bibitem{rebel2020}
Noam Brown, Anton Bakhtin, Adam Lerer, and Qucheng Gong.
\newblock Combining deep reinforcement learning and search for
imperfect-information games.
\newblock \emph{Advances in Neural Information Processing Systems}, 2020.
\newblock \url{https://proceedings.neurips.cc/paper/2020/hash/c61f571dbd2fb949d3fe5ae1608dd48b-Abstract.html}

\end{thebibliography}

\end{document}
