\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{tabularx}

\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  citecolor=blue,
  urlcolor=blue
}
\setlength{\emergencystretch}{3em}
\allowdisplaybreaks

\newtheorem{assumption}{Assumption}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{observation}{Observation}

\newcommand{\stl}{\textsc{STL}}
\newcommand{\hal}{\mathrm{Hal}}
\newcommand{\baku}{\mathrm{Baku}}
\newcommand{\states}{\mathcal{S}}
\newcommand{\actions}{\mathcal{A}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\prob}{\mathbb{P}}
\newcommand{\valuefn}{V}
\newcommand{\simplex}{\Delta}
\newcommand{\val}{\operatorname{val}}
\newcommand{\gap}{\operatorname{gap}}

\title{From Exact Anchors to Simultaneous-Move AlphaZero\\
for the \stl{} Handkerchief Game}
\author{Repository whitepaper}
\date{July 10, 2026}

\begin{document}
\maketitle

\begin{abstract}
This whitepaper records both the implemented solver and the revised plan for
reinforcement learning. The game is modeled as a fully observed, two-player,
zero-sum stochastic game with simultaneous exact-second actions. The Python
engine is the rule oracle. The rigorous solver performs finite-horizon
backward induction with engine-derived chance and linear-programming minimax
at every simultaneous node. The approximate stack provides candidate search,
matrix-game Monte Carlo tree search (MCTS), tablebase frontiers, and a
policy/value network.

The repository does not yet implement a complete AlphaZero loop. Phases P0--P3
of the Python bridge now provide a finite episode contract, exact/search
conformance, reconstructable replay and checkpoints, and a tested
simultaneous-MCTS improvement object; a short Generation-Zero pipeline smoke
also passes. Full anchor generation is in progress and full training remains
unexecuted. The compatibility self-play module still uses a legacy bucketed agent against
scripted opponents. This distinction matters: AlphaZero-style reinforcement
learning requires search policies at states actually visited by MCTS self-play
and terminal episode outcomes as value targets. Rust is deferred until the
Python behavior and evidence package are stable. We prove the finite-horizon exact result, the
non-expansiveness of minimax backup, a frontier-error bound, and a local saddle
gap bound. We also state precisely why these results do not prove convergence
or global exploitability of the learned agent.
\end{abstract}

\section{Scope, Status, and Claims}

The live runtime tree is under \path{stl/}. The disconnected old package mirror
was removed after pushed archival commit \texttt{a4b03db}; Git, rather than a
second source tree, retains that history. The implementation plan is
\path{docs/REGEN2RL.md}, and the granular removal ledger is
\path{docs/DISTILLATION_AUDIT.md}. This paper supplies the mathematical
rationale and claim boundary.

\begin{center}
\begin{tabularx}{\textwidth}{@{}lX@{}}
\toprule
Path & Responsibility \\
\midrule
\path{stl/engine/} & Rule authority, clock, roles, legality, death, revival, and leap behavior \\
\path{stl/solver/exact.py} & Exact public state, engine expansion, LP minimax, and finite-horizon solve \\
\path{stl/solver/search.py} & Candidate search, bounded resolve, leaf adapters, and matrix-game MCTS \\
\path{stl/solver/tablebase.py} & Tactical fixtures, certified intervals, and Tier A lookup \\
\path{stl/learning/} & Targets, policy/value model, training, replay support, and gates \\
\path{stl/play/} & Playable agents, scripted opponents, environment wrappers, and legacy bucket baseline \\
\path{stl/commands/} & Hydra command implementations \\
\path{configs/} & Resolved experiment and command parameters \\
\bottomrule
\end{tabularx}
\end{center}

The following claims are justified now:

\begin{itemize}[leftmargin=*]
  \item engine-relative correctness of exact action legality and transition
        expansion, subject to the regression suite;
  \item exact minimax value for the finite-horizon game actually enumerated by
        the full-width LP solver;
  \item exact values of pinned analytic fixtures by their construction; and
  \item finite-depth best-response intervals under their reported frontier
        brackets.
\end{itemize}

The following claims are not justified now:

\begin{itemize}[leftmargin=*]
  \item that the unrestricted-duration game has been solved;
  \item that finite-budget candidate MCTS converges under the present
        selection rule;
  \item that held-out mean squared error implies low exploitability;
  \item that existing self-play code is AlphaZero self-play; or
  \item that a Rust implementation is required for, or already defines, the
        learning baseline.
\end{itemize}

\section{Game Model}

\subsection{Fully Observed State with Simultaneous Actions}

Let the players be \(\mathcal{N}=\{\hal,\baku\}\). At a nonterminal public
state \(s\in\states\), both players observe the complete state represented by
\path{ExactPublicState}: player cylinders, total time dead, death counts, life
status, physicality, referee CPR count, clock, half-round, round number, first
dropper, and terminal fields. The players then choose their current actions
without observing the other current action. This is a perfect-state or fully
observed simultaneous-move game, not an alternating-action board game and not
a hidden-state poker game.

The engine assigns a dropper \(D(s)\) and checker \(C(s)\). It is convenient
to write the two player action sets as \(A_{\hal}(s)\) and
\(A_{\baku}(s)\), then map the resulting pair to a drop second \(d\) and a
check second \(c\) according to the roles. The code stores matrices as
drop-by-check and transposes or negates them so that Hal remains the maximizing
player.

\subsection{Actor-Aware Exact-Second Legality}

Normal legal seconds are \(1,\ldots,60\). In the leap window only Baku, when
Baku is the dropper, may select second 61. Checkers are always capped at 60,
and Hal can never select 61. Thus
\[
A_D(s)=
\begin{cases}
\{1,\ldots,61\}, & D(s)=\baku \text{ and }s\text{ is a leap turn},\\
\{1,\ldots,60\}, & \text{otherwise},
\end{cases}
\qquad
A_C(s)=\{1,\ldots,60\}.
\]
Dense learned policies have length 62. Index 0 is illegal padding and index
\(j\) denotes literal second \(j\).

The check succeeds iff \(c\ge d\). On success the checker accumulates
\(c-d\) seconds, including zero on the diagonal. On failure the checker incurs
the fixed 60-second penalty. All death, revival, clock, and leap effects then
come from \path{stl/engine/game.py}.

\subsection{Transition Kernel and Utility}

Let
\[
P(s'\mid s,a_{\hal},a_{\baku})
\]
be the transition kernel induced by the engine, including survival and fatal
branches. The rigorous solver does not learn or reimplement this kernel.

Terminal utility is always Hal-perspective:
\[
u(s)=
\begin{cases}
+1, & s\text{ is terminal and Hal wins},\\
-1, & s\text{ is terminal and Baku wins},\\
0, & s\text{ is an explicitly declared draw}.
\end{cases}
\]
There is no intermediate shaped reward in \path{stl/solver}.

\subsection{Why Reinforcement Learning Needs an Episode Contract}

\begin{observation}[The engine does not force a finite episode]
A repeated diagonal action pair \(c=d\) succeeds with squandered time zero.
Such a path need not increase either cylinder and can therefore remain
nonterminal for arbitrarily many half-rounds.
\end{observation}

Consequently, a completed-game outcome is not guaranteed merely by calling
the engine in a loop. The first Python RL baseline defines a capped training
game \(G_H\): after \(H\) half-rounds, a nonterminal trajectory is assigned
training result zero and marked as truncated. The initial plan freezes
\(H_{\mathrm{train}}=64\) and \(H_{\mathrm{eval}}=128\), then gates promotion
on low truncation rate and sensitivity to the cap. This is an RL-wrapper
adjudication, not a new engine terminal rule. Any result learned under this
contract is a result about \(G_H\), not automatically about the
unrestricted-duration game.

\section{Exact Finite-Horizon Solver}

\begin{assumption}[Engine authority]
The Python engine in \path{stl/engine} is the rule oracle. Solver correctness
is relative to its legality and transition functions.
\end{assumption}

Let \(b:\states\to[-1,1]\) be a boundary evaluator. For a terminal state,
the boundary is ignored in favor of \(u\). Define
\[
V_0^b(s)=
\begin{cases}
u(s), & s\text{ terminal},\\
b(s), & s\text{ nonterminal},
\end{cases}
\]
and, for \(h>0\), define the Hal-payoff matrix
\[
M_h^b(s)_{ij}
=
\sum_{s'}P(s'\mid s,a_{\hal,i},a_{\baku,j})
V_{h-1}^b(s').
\]
The Bellman-Shapley backup is
\[
V_h^b(s)=\val\!\left(M_h^b(s)\right)
=
\max_{\sigma\in\simplex(A_{\hal}(s))}
\min_{\tau\in\simplex(A_{\baku}(s))}
\sigma^{\mathsf T}M_h^b(s)\tau.
\]

For a matrix \(M\in\mathbb{R}^{m\times n}\), the row-player LP is
\[
\begin{aligned}
\max_{\sigma,v}\quad &v\\
\text{subject to}\quad
&\sum_{i=1}^{m}\sigma_i M_{ij}\ge v,
&&j=1,\ldots,n,\\
&\sum_{i=1}^{m}\sigma_i=1,\\
&\sigma_i\ge0,
&&i=1,\ldots,m.
\end{aligned}
\]

The certified code path uses full legal exact-second actions and this LP. With
the current terminal-only exact cutoff, \(b(s)=0\) and unresolved probability
is tracked separately.

\begin{proposition}[Exact finite-horizon value]
For fixed engine state \(s\), finite horizon \(h\), full legal action sets,
and boundary evaluator \(b\), recursive engine expansion followed by exact LP
minimax returns \(V_h^b(s)\).
\end{proposition}

\begin{proof}
The claim is immediate for \(h=0\). Assume it holds at \(h-1\). For each
legal joint action at \(s\), the engine enumerates all successor chance
branches and the induction hypothesis gives the continuation value at every
successor. Their probability-weighted sum is therefore the corresponding
entry of \(M_h^b(s)\). The one-step decision is a finite zero-sum matrix game,
so von Neumann's minimax theorem equates its LP value with
\(\val(M_h^b(s))\). This is \(V_h^b(s)\), completing the induction.
\end{proof}

\section{Error Reasoning for Depth-Limited Search}

\subsection{Minimax Backup Is Non-Expansive}

\begin{lemma}[Matrix-value Lipschitz bound]
If two equally shaped payoff matrices satisfy
\(\lVert M-\widehat M\rVert_{\infty}\le\epsilon\), then
\[
\left|\val(M)-\val(\widehat M)\right|\le\epsilon.
\]
\end{lemma}

\begin{proof}
For every pair of mixed strategies \((\sigma,\tau)\),
\[
\left|\sigma^{\mathsf T}(M-\widehat M)\tau\right|
\le\epsilon
\]
because \(\sigma\) and \(\tau\) are probability vectors. Taking a minimum
and then a maximum cannot enlarge this uniform bound. Applying the same
argument with \(M\) and \(\widehat M\) exchanged gives the absolute-value
claim.
\end{proof}

\begin{proposition}[Frontier error does not amplify under exact backup]
Suppose two boundary evaluators satisfy
\(\sup_s|b(s)-\widehat b(s)|\le\epsilon\). If all actions and transitions are
expanded exactly, then for every finite horizon \(h\),
\[
\sup_s|V_h^b(s)-V_h^{\widehat b}(s)|\le\epsilon.
\]
\end{proposition}

\begin{proof}
At depth zero the result is the hypothesis. If it holds at depth \(h-1\),
then every action-cell continuation differs by at most \(\epsilon\), since an
expectation is non-expansive in the supremum norm. The matrix-value lemma then
bounds the backed-up value difference by \(\epsilon\). Induction proves the
claim.
\end{proof}

This proposition justifies calibrated frontier values for an exact
depth-limited resolve. It does not say that a finite-sample MCTS estimate has a
uniform error bound, nor that a compact feature map can represent such a
bound.

\subsection{Local Policy Quality from Matrix Accuracy}

For a Hal row strategy \(\widehat\sigma\) and Baku column strategy
\(\widehat\tau\), define the exact matrix saddle gap
\[
\gap_M(\widehat\sigma,\widehat\tau)
=
\max_{\sigma}\sigma^{\mathsf T}M\widehat\tau
-
\min_{\tau}\widehat\sigma^{\mathsf T}M\tau.
\]
It is zero at an exact saddle point and measures the total one-step gain
available from unilateral deviations.

\begin{proposition}[Approximate-matrix saddle gap]
Let \((\widehat\sigma,\widehat\tau)\) be an exact equilibrium of
\(\widehat M\). If
\(\lVert M-\widehat M\rVert_{\infty}\le\epsilon\), then
\[
\gap_M(\widehat\sigma,\widehat\tau)\le2\epsilon.
\]
\end{proposition}

\begin{proof}
Let \(\widehat v=\val(\widehat M)\). Uniform matrix error implies
\[
\max_{\sigma}\sigma^{\mathsf T}M\widehat\tau
\le \widehat v+\epsilon,
\qquad
\min_{\tau}\widehat\sigma^{\mathsf T}M\tau
\ge \widehat v-\epsilon.
\]
Subtracting the second inequality from the first gives the result.
\end{proof}

This provides a concrete policy gate: whenever full-width exact matrices are
available, report saddle gap in addition to value error.

\subsection{Candidate Sets Need Their Own Audit}

No analogous bound follows merely because a candidate set contains many
actions or tactically plausible seconds. A restricted matrix can be identically
zero while one omitted Hal action pays \(+1\) against every Baku action. The
restricted value is then zero and the full value is one. Candidate search is
therefore an approximation whose omitted-action best-response gain must be
measured against full width on a frozen audit distribution.

\section{Current Search Stack}

\subsection{Selective Solve and Bounded Resolve}

\path{stl/solver/search.py} generates candidate seconds around endpoints,
diagonals, safe budgets, and overflow boundaries. Its selective recursion
still uses engine transitions and role-correct matrix minimax, but its action
restriction is approximate. Critical-state resolve applies a hard horizon,
usually one half-round, and evaluates the frontier with a tablebase or network
adapter. LP remains the certification path; Python CFR+ is a runtime option for
bounded local matrices.

\subsection{Matrix-Game MCTS}

Each MCTS node stores candidate role actions, a Hal-perspective mean
\(Q\)-matrix, joint cell counts, priors, children, and an exact engine
snapshot. The current exploration term is
\[
U_{ij}=c_{\mathrm{puct}}P_{ij}
\frac{\sqrt{N}}{1+N_{ij}}.
\]
The Hal maximizer solves an optimistic matrix \(Q+U\); the opponent solves an
optimistic matrix for its own payoff, equivalently \(-Q+U\). Joint actions are
sampled from the two resulting role marginals, chance is sampled through the
engine, and Hal-perspective leaf values are backed up without alternating sign.

Let \(\widehat\sigma_D^{(t)}\) and
\(\widehat\sigma_C^{(t)}\) be role equilibria of the root's empirical
mean-\(Q\) matrix before the \(t\)-th root selection. The canonical improvement
policy is their linearly weighted average
\[
\pi_D=\frac{1}{W_T}\sum_{t=1}^{T}t\widehat\sigma_D^{(t)},
\qquad
\pi_C=\frac{1}{W_T}\sum_{t=1}^{T}t\widehat\sigma_C^{(t)},
\qquad
W_T=\sum_{t=1}^{T}t.
\]
Later empirical matrices receive greater weight without collapsing mixed play.
The optimism-augmented equilibria select exploratory joint cells but are not
reported as policy probability. Network priors are masked and normalized per
role, their outer product defines the joint prior, and optional root Dirichlet
noise perturbs the two legal marginals independently. The final saddle point
of the mean \(Q\)-matrix remains a separate value/policy diagnostic; in a
nearly flat matrix an LP may return an arbitrary pure vertex, so it is not the
canonical training or deployed policy.

The convergence theorem for simultaneous-move MCTS in
\cite{lisy2013} assumes Hannan-consistent selection, guaranteed exploration,
and a finite game tree. The repository's PUCT-like matrix selection, neural
frontier, candidate actions, transpositions, and finite episode wrapper are not
automatically covered by that theorem. The frozen exact-matrix convergence pack
therefore supplies empirical, not theorem-level, evidence: across 120
fixture/budget/seed records its 1024-iteration median absolute value error is
zero, 95th-percentile error is \(0.001812\), maximum saddle gap is
\(0.005726\), maximum per-fixture root-value standard deviation is
\(0.000736\), and no fixture worsens from 256 to 1024 iterations. A future
regret-matching SM-MCTS replacement remains appropriate if broader frozen
packs violate the declared gate.

\section{Audit of the Current Learning Attempt}

\subsection{Implemented Pieces}

The repository already contains several components needed by the target
system:

\begin{itemize}[leftmargin=*]
  \item \path{stl/learning/model.py}: a shared trunk, Hal value head, and two
        length-62 role-policy heads driven by 52 named V2 features;
  \item \path{stl/learning/contracts.py}: the versioned finite-episode,
        perspective, configuration, and horizon-sensitivity contract;
  \item \path{stl/learning/replay.py}: reconstructable
        \path{TrainingRecordV2} rows, safe hash-verified shards, collision
        audits, and grouped splits;
  \item \path{stl/learning/train.py}: value MSE, masked role-policy cross
        entropy, source weights, grouped selection, and strict checkpoint
        bundles;
  \item \path{stl/learning/targets.py}: exact/tablebase labels and MCTS
        bootstrap labels;
  \item \path{stl/learning/support/reanalysis.py}: deeper exact attempts and a
        high-iteration MCTS fallback;
  \item calibration, audit, ladder, policy-drift, finite-depth best-response,
        SPRT, and promotion utilities; and
  \item \path{stl/play/agent.py}: a playable network-plus-MCTS agent that
        samples the canonical improved role policies.
\end{itemize}

The P1 and P3 frozen conformance packs test exact transition/orientation
parity, omitted-action gains, chance sampling, and horizon-one MCTS convergence.
The P4 smoke writes and reloads small Generation-Zero V2 train and external
ruler shards. These are valuable prerequisites; they do not yet form one
closed RL loop.

\subsection{Blocking Integration Gaps}

\begin{itemize}[leftmargin=*]
  \item \textbf{No premature command surface.}
        The broken legacy \path{command=self_play} wrapper has been removed.
        A public command is added only when P5 supplies the joint MCTS actor,
        reconstructable replay, and command-contract tests.
  \item \textbf{Legacy actor.}
        \path{stl/learning/support/self_play.py} uses bucketed
        \path{CanonicalHal} against scripted opponents. It is curriculum or
        legacy outcome data, not MCTS self-play by both sides.
  \item \textbf{Expert-iteration labels.}
        MCTS bootstrap sweeps a constructed state grid and labels root value;
        it does not supply trajectory outcomes from the improved policy.
  \item \textbf{Generation Zero is not frozen.}
        The structural smoke is complete, but the full exact/tablebase anchor
        corpus, external-ruler review, training run, and search-in-the-loop
        gate have not run.
  \item \textbf{Incomplete promotion evidence.}
        Policy drift is optional in promotion, and finite-depth best response
        is called certified exploitability; the first omits a policy gate and
        the second can be misread as a global claim.
\end{itemize}

The correct response is not to discard the exact stack. It is to connect the
existing pieces with explicit schemas and gates, while quarantining legacy
bucket search from the new self-play actor.

\section{Target Python AlphaZero System}

\subsection{Reconstructable State and Replay}

Each \path{TrainingRecordV2} row carries a versioned 52-field feature vector
and a serialized exact public state. It also records both legal masks, both
improved role policies, Hal-perspective value target, target kind, episode and
half-round indices, truncation status, parent checkpoint digest, search/config
digest, schema versions, and RNG seeds. Pickle-free numeric shards and a
canonical JSON manifest are written atomically and verified by hashes. This
turns replay into an auditable experiment record rather than an opaque tensor
dump.

The V2 encoder includes every nonconstant value-relevant field in
\path{ExactPublicState} and appends documented relative features. A
compact network remains an approximation, but replay reconstruction permits
collision analysis and exact or MCTS reanalysis. Split assignment is grouped
by exact-state hash and episode so repeated or replicated states cannot cross
train and validation boundaries.

\subsection{Policy/Value Function}

The network is
\[
f_{\theta}(x(s))=
\left(p_D(\cdot\mid s),p_C(\cdot\mid s),v_{\theta}(s)\right),
\]
where each policy is masked and normalized over the actor-aware legal seconds
for its role, and \(v_{\theta}(s)\in[-1,1]\) estimates Hal outcome. Role heads
are appropriate because the state encoder identifies which named player holds
each role.

The two policies define independent mixed actions. A joint correlated policy
would define a different strategic object, because it could coordinate the
players' simultaneous actions. Therefore the joint prior and sampling law are
factorized:
\[
P_{ij}=p_D(i\mid s)p_C(j\mid s),
\qquad
d\sim\pi_D(\cdot\mid s),
\quad
c\sim\pi_C(\cdot\mid s).
\]

\subsection{MCTS Self-Play}

At every visited state, one search produces both improved role marginals.
Self-play exploration perturbs the legal marginals separately and records the
noise seed. Evaluation disables noise. Because a simultaneous equilibrium may
require mixing, evaluation samples the noise-free marginals rather than taking
an argmax.

The actor stores temporary rows
\[
(s_t,\pi_{D,t},\pi_{C,t},m_{D,t},m_{C,t})
\]
and samples the next engine transition. At the end of the episode, it attaches
\(z\in\{-1,0,+1\}\) to every row, with zero reserved for the explicit capped
draw contract when no engine draw exists. Root MCTS value is logged for
diagnosis but is not mislabeled as the terminal outcome.

\subsection{Learning Objective}

For replay sample \((s,\pi_D,\pi_C,z)\), the AlphaZero-style loss adapted to
two role heads is
\[
\mathcal{L}(\theta)=
\lambda_v\left(z-v_{\theta}(s)\right)^2
-\lambda_D\sum_{a\in A_D(s)}\pi_D(a\mid s)\log p_D(a\mid s)
-\lambda_C\sum_{a\in A_C(s)}\pi_C(a\mid s)\log p_C(a\mid s)
+c\lVert\theta\rVert_2^2.
\]
This is the direct analogue of the AlphaZero objective
\cite{alphazero2017}, with separate simultaneous role policies. Exact and
tablebase anchors may substitute certified value and equilibrium-policy
targets. Rows without a certified policy have that policy loss disabled; they
do not receive a fabricated uniform label.

The initial replay mixture keeps 25 percent immutable exact/tablebase anchors
and 75 percent self-play rows. This ratio is an engineering baseline, not a
convergence theorem. Its purpose is to prevent catastrophic drift from exact
anchors while the empirical promotion gates judge whether self-play improves
strength.

\subsection{Reanalysis}

Reanalysis reconstructs stored states and refreshes the search policies using
the current frozen champion. Completed episode outcomes remain unchanged.
Exact and tablebase rows are immutable. Search-bootstrap values, if retained
for experiments, carry a distinct target-kind tag so they cannot be confused
with Monte Carlo outcomes.

\section{Promotion and Evidence}

Promotion is more conservative than AlphaZero's policy of continually using
the latest network. AlphaGo Zero's evaluate-and-select lineage, together with
the repository's observed calibration and strength regressions, motivates a
frozen champion/candidate gate.

Every candidate must pass:

\begin{enumerate}[leftmargin=*]
  \item artifact and configuration digest consistency;
  \item exact/tablebase value calibration and role-policy legality;
  \item the exact-matrix MCTS convergence pack;
  \item a paired two-direction arena that respects Hal/Baku asymmetry and uses
        a predeclared SPRT;
  \item non-regression against the scripted robustness league, including
        \path{pattern_reader};
  \item finite-depth best-response interval comparison at identical scenarios,
        horizons, support mass, and frontiers; and
  \item a mandatory policy-drift/readability report with exact-anchor saddle
        gaps.
\end{enumerate}

Best-response intervals certify only their bounded calculation. If
\([L_C,U_C]\) and \([L_B,U_B]\) overlap, the comparison is inconclusive. A
promotion report may say that no sampled case is certified worse; it may not
infer global improvement or global low exploitability.

\section{Phased Implementation and Rust Boundary}

The executable plan and numeric exit criteria live in
\path{docs/REGEN2RL.md}. Its dependency chain is:

\begin{enumerate}[leftmargin=*]
  \item freeze the objective, perspective, and finite episode contract;
  \item certify engine/exact/search conformance in Python;
  \item introduce reconstructable replay and checkpoint schemas;
  \item validate matrix-game MCTS as the improvement operator;
  \item train a supervised exact-anchor generation zero;
  \item build genuine MCTS self-play for both roles;
  \item close replay, training, and reanalysis;
  \item integrate statistical promotion; and
  \item freeze a reproducible Python baseline and evidence package.
\end{enumerate}

Items 1--4 (P0--P3) are implemented and verified. A short item-5 pipeline
smoke generated 26 reconstructable training rows and 26 external-ruler rows
and reloaded a one-epoch checkpoint. The full item-5 anchor generation is the
current stopping boundary; neither that long job nor full training has run.

No phase in this chain requires new Rust work. Existing Rust CFR+ bindings are
opt-in comparison fixtures only. After the Python baseline is frozen, narrow
profiled kernels may be ported behind parity tests. Python remains the oracle
for legal actions, transitions, policy normalization, replay semantics, and
promotion. The Rust acceptance ordering is behavioral parity first and speed
second.

\section{What the Proofs Do and Do Not Establish}

The exact induction proves a finite-horizon result when all legal actions and
chance branches are represented and the LP is solved. The non-expansive backup
and saddle-gap propositions explain how a genuine uniform frontier or matrix
error bound would control local value and policy error.

They do not provide such a uniform error bound for a neural network. Empirical
MSE is distribution-dependent; feature collisions can force incompatible
states to share a prediction. Finite-budget MCTS adds sampling and selection
error. Candidate sets add unmeasured action error unless full-width audits are
performed. Function approximation, replay nonstationarity, and policy/value
co-adaptation remove the assumptions of tabular exact policy iteration.

Accordingly, the planned system is a rigorously anchored empirical solver, not
a theorem of global AlphaZero convergence. A future global claim would require
at least a well-defined unrestricted objective, exhaustive or otherwise valid
full-game bounds, and a global exploitability certificate. Until then, reports
must keep these categories separate:

\begin{itemize}[leftmargin=*]
  \item \textbf{exact}: full-width finite-horizon LP or analytic tablebase;
  \item \textbf{bounded}: interval or error statement for a declared horizon
        and frontier;
  \item \textbf{approximate}: finite-budget candidate MCTS or neural output;
  \item \textbf{empirical}: self-play, arena, calibration, and robustness
        results on declared distributions.
\end{itemize}

\section{Validation Protocol}

The minimum documentation-aligned regression commands are:

\begin{verbatim}
uv run pytest tests/engine -q
uv run pytest tests/solver -q
uv run pytest tests/learning -q
uv run pytest tests/play -q
uv run pytest -q
uv run pytest --collect-only -q
\end{verbatim}

The stable-baseline gate additionally requires deterministic self-play replay,
exact-state reconstruction, grouped split isolation, exact-matrix MCTS budget
curves, a successful and a rejected promotion dry-run, horizon sensitivity,
and a graph refresh by running \texttt{graphify update} against the repository
root; generated
checkpoints and reports remain outside source review unless the artifact itself
is under audit.

\section{Conclusion}

The repository has a credible exact/search foundation: engine authority,
literal-second stochastic minimax, tablebase anchors, bounded resolve, a tested
matrix-game MCTS improvement object, reconstructable replay/checkpoints, and a
policy/value learner. The remaining gap is a frozen Generation Zero, genuine
MCTS self-play trajectories with terminal outcomes, a closed replay/reanalysis
loop, and mandatory statistical promotion. Only after those Python phases
produce a stable evidence package is the baseline suitable as an oracle for a
later Rust kernel.

\begin{thebibliography}{99}

\bibitem{vonneumann1928}
John von Neumann.
\newblock Zur Theorie der Gesellschaftsspiele.
\newblock \emph{Mathematische Annalen}, 100:295--320, 1928.
\newblock \url{https://doi.org/10.1007/BF01448847}

\bibitem{shapley1953}
Lloyd S. Shapley.
\newblock Stochastic games.
\newblock \emph{Proceedings of the National Academy of Sciences},
39(10):1095--1100, 1953.
\newblock \url{https://doi.org/10.1073/pnas.39.10.1095}

\bibitem{zinkevich2007}
Martin Zinkevich, Michael Johanson, Michael Bowling, and Carmelo Piccione.
\newblock Regret minimization in games with incomplete information.
\newblock \emph{Advances in Neural Information Processing Systems}, 2007.
\newblock \url{https://papers.nips.cc/paper_files/paper/2007/hash/08d98638c6fcd194a4b1e6992063e944-Abstract.html}

\bibitem{lisy2013}
Viliam Lisy, Vojtech Kovarik, Marc Lanctot, and Branislav Bosansky.
\newblock Convergence of Monte Carlo tree search in simultaneous move games.
\newblock arXiv:1310.8613, 2013.
\newblock \url{https://arxiv.org/abs/1310.8613}

\bibitem{deepstack2017}
Matej Moravcik, Martin Schmid, Neil Burch, Viliam Lisy, Dustin Morrill,
Nolan Bard, Trevor Davis, Kevin Waugh, Michael Johanson, and Michael Bowling.
\newblock DeepStack: Expert-level artificial intelligence in heads-up
no-limit poker.
\newblock \emph{Science}, 356(6337):508--513, 2017.
\newblock \url{https://doi.org/10.1126/science.aam6960}

\bibitem{alphagozero2017}
David Silver, Julian Schrittwieser, Karen Simonyan, Ioannis Antonoglou,
Aja Huang, Arthur Guez, Thomas Hubert, Lucas Baker, Matthew Lai, Adrian Bolton,
Yutian Chen, Timothy Lillicrap, Fan Hui, Laurent Sifre,
George van den Driessche, Thore Graepel, and Demis Hassabis.
\newblock Mastering the game of Go without human knowledge.
\newblock \emph{Nature}, 550:354--359, 2017.
\newblock \url{https://doi.org/10.1038/nature24270}

\bibitem{alphazero2017}
David Silver, Thomas Hubert, Julian Schrittwieser, Ioannis Antonoglou,
Matthew Lai, Arthur Guez, Marc Lanctot, Laurent Sifre, Dharshan Kumaran,
Thore Graepel, Timothy Lillicrap, Karen Simonyan, and Demis Hassabis.
\newblock Mastering chess and shogi by self-play with a general reinforcement
learning algorithm.
\newblock arXiv:1712.01815, 2017.
\newblock \url{https://arxiv.org/abs/1712.01815}

\bibitem{rebel2020}
Noam Brown, Anton Bakhtin, Adam Lerer, and Qucheng Gong.
\newblock Combining deep reinforcement learning and search for
imperfect-information games.
\newblock \emph{Advances in Neural Information Processing Systems}, 2020.
\newblock \url{https://proceedings.neurips.cc/paper/2020/hash/c61f571dbd2fb949d3fe5ae1608dd48b-Abstract.html}

\end{thebibliography}

\end{document}
